---
title: "How to use the function check_last_successful_run"
output:
  rmarkdown::html_vignette:
  tabset: TRUE
vignette: >
  %\VignetteIndexEntry{How to use check_last_successful_run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r eval, include=FALSE}
darautils_installed <- rlang::is_installed("DARAutils")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r lintstart, include FALSE}
# nolint start: object_usage_linter
```

First, we need to load the required packages.
```{r setup, include=TRUE, eval=FALSE}
library(DARApipeline)
```

```{r lintend, include FALSE}
# nolint end: object_usage_linter
```

### Introduction

The `check_last_successful_run()` function has been developed to quickly determine the last successful production run. Not only is the user able to check the current project, but every project (provided they have the path to the project).


The parameters, and basic examples can be found in the documentation of the function. This can be found in the references tab, or in R using:

```{r help, eval=FALSE}
?check_last_successful_run
```

### Basic example
To use this function, the user simply, only has to provide a file path to the log directory of the project they want to check. As seen below:

```{r basic_example, eval=FALSE}
check_last_successful_run(log_dir = "<full_path_to_project_on_production>/logs")

```

#### Automatic search
We can imagine that you might not always know the path to the project you want to check. In that case, you can make a minor addition to the `file_paths.yaml`. With this addition, the user only has to call the function `check_last_successful_run()`, which automatically searches the production log files for the current project (even when their working from their home project)!.

> Note: To use the automatic search, the user must have initialized their pipeline. This can be done with the `pipeline_init()` function, or running their `prepare.R` script.

The minor tweak means adding the following line to the `file_paths.yaml`: <br>
`dir_mnt: /rivm/EPI` <br>

The updated `file_paths.yaml` should look like this:

```{yaml example, eval=FALSE}
dir_outputs: your_output_dir_path
dir_scripts: your_scripts_dir_path
dir_project: your_project_dir_path
dir_mnt: "/rivm/EPI"

```

> Note: The updated `file_paths.yaml` can also be pushed and merged, meaning others users don't have to update their config files.
> Newly generated projects with cookiecutter will already have this updated config.

### Custom parameters
The user can also specify custom parameters, these will be explained below.

#### Include cronjob
With this parameter, the user can specify if you want to include cronjobs or not. Default is FALSE
*This might slow down the search, as it will have to search through all cronjob logs.*
```{r cronjob_example, eval=FALSE}
check_successful_run(log_dir = "<full_path_to_project_on_production>/logs",
                     include_cronjob = TRUE)
```

#### Cronjob only
With this parameter, the user can specify to only search for cronjobs logs files. Default is FALSE.
When set to TRUE, it overrides the `include_cronjob` parameter.
```{r cronjob_only_example, eval=FALSE}
check_successful_run(log_dir = "<full_path_to_project_on_production>/logs",
                     cronjob_only = TRUE)
```


#### Search window
With this parameter, the user can specify a search window. This means that the function will only search for successful runs within the last number of specified lines. Default is 1 line.

This is especially useful when the completion messsage isn't on the last line (perhaps due to warnings, or other messages).
```{r search_window_example, eval=FALSE}
check_successful_run(log_dir = "<full_path_to_project_on_production>/logs",
                     search_window = 5)
```

#### Search line
With this parameter, the user can specify custom completion lines. Currently, the standard completion lines are: <br>
*"=== Save step done! ==="* <br>
*"=== Reporting step done! ==="* <br>

Different completion lines should be specified in the `search_line` parameter.

```{r search_line_example, eval=FALSE}
check_last_successful_run(log_dir = "<full_path_to_project_on_production>/logs",
                          search_line = "Infectieradar saved clean joined, weekly and intake data.")
```
