---
title: "DARA EPIpipeline and config introduction"
output:
  rmarkdown::html_vignette:
  tabset: TRUE
vignette: >
  %\VignetteIndexEntry{Pipeline design and config}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r eval, include=FALSE}
DARApipeline_installed <- rlang::is_installed("DARApipeline") # nolint: namespace_linter
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r lintstart, include=FALSE}
# nolint start: object_usage_linter
```

First, we need to load the required packages.
```{r setup, include=TRUE, eval=TRUE}
library(DARApipeline)
```

```{r lintend, include=FALSE}
# nolint end: object_usage_linter
```

# Introduction

**DARApipeline** functions are designed to help epidemiologists set up their infectious disease surveillance data pipelines.
In this article, we will explain how the `config.yaml` files are used and how they interact with the `DARApipeline` functions. The main goal is to eliminate repeated code for loading input, deciding what to run, and saving output. At the same time, we aim to achieve a shared project structure across different surveillance projects.
We refer to figures, tables, and maps collectively as *objects*. The pipeline:

- loads all input required for each object
- determines what to run based on dependencies between objects, tags or explicit calls
- saves each object in the specified output format

The parameters and basic examples can be found in the function documentation, either via the References tab or directly in R using:

```{r help, eval=FALSE}
?pipeline_run
```

## initiating the pipeline
`pipeline_init()` is the starting function of the `DARApipeline`. All other `pipeline_` functions depend on it and will not work if this function has not been run. It performs a few important steps:

1. Creates a `run_timestamp` (date + hour + minute) at the moment `pipeline_init()` is executed.
2. Initiates logging in the `/logs` folder.
3. Reads and interprets the configuration files in `config/`.

There are some additional, smaller background features which we will skip for now.

```{r helper, eval=TRUE, include=FALSE}
library(DARAutils)
local_logger_sink()
withr::local_options(list(
  DARApipeline.configdir = file.path("../tests/testthat/fixtures/configs/config"),
  DARApipeline.skiplogging = TRUE
))
```

```{r init, eval=TRUE}
pipeline_init()
```

# config files

As mentioned, `pipeline_init()` reads the configuration files. There are three essential YAML configuration files:

- `config/base/file_paths.yaml`
- `config/base/object_definitions.yaml`
- `config/base/object_relations.yaml`

Further, there is one extra, optional, YAML configuration file:
- `config/base/object_schedules.yaml`

In general, for YAML files, names between curly brackets indicate variables, e.g. `{RUN_TIMESTAMP}`. This means that the value of `RUN_TIMESTAMP`, created by the `pipeline_init()` function, is inserted at that position.

We use this mechanism to store logs, cache, and output by date-time.

## config file `object_definitions.yaml`

This config file contains all object-specific information, except the relations between objects and schedules of objects.

The first key is the `__defaults__` key. These are the default values applied when an object is missing a specific key. This helps reduce duplication (for example, avoiding a `skip_cache: no` line for every object).

`type` indicates what type the entry is. We currently use three types:

1. `type: object` – a table, figure, map, or HTML object created in this pipeline.
2. `type: data` – a data `.rds` coming from a datasource pipeline.
3. `type: other` – a fixed data file to import. This can for example be a static csv file.


#### Mapping keys for `type: object`

These are the possible mapping keys for objects of `type: object`:

- `skip_cache`
  By default, objects are cached, i.e. intermediately saved. The `skip_cache` key
  indicates whether caching should be skipped.
  Allowed values: `yes` or `no`.

- `tag`
  The tag name for this object. Tags are used to create logical groups of objects.
  Groups of objects with the same tag can be defined as follows:

```{yaml example_tags, eval=FALSE}
my_object1:
  type: object
  tag: tag_a
my_object2:
  type: object
  tag: tag_a
my_object3:
  type: object
  tag: tag_b
```

Objects with the same tag can then be run as follows:

```{r  eval = FALSE}
# This example creates object1 and object2 and not object 3
pipeline_run(tags = "tag_a")
```

- `output_formats`
The saving format(s) of the object. Use the file extension names. Default formats include: csv, png, svg, and xlsx.

```{yaml example_output_formats, eval=FALSE}
my_object1:
  type: object
  output_formats: xlsx
```


- `output_arguments`
The arguments which are passed to the output saving functions.
```{yaml example_output_arguments, eval=FALSE}
my_table:
  type: object
  output_formats:
    - csv
  output_arguments:
    csv:
      na: ""
      row.names: FALSE
```

- `output_formats_custom`
Custom file extensions and saving functions when you want to use a different format than those available in output_formats.
It expects a mapping like:
```{yaml example_output_formats_custom, eval=FALSE}
my_object:
  output_formats_custom:
    rds: "saveRDS"
```

- `cache_dir`
Directory where the cache should be saved if `skip_cache: no`.
The value should be a directory that varies per run, for example:
```{yaml example_cache_dir, eval=FALSE}
my_object:
  cache_dir: cache/{RUN_TIMESTAMP}
```

#### HTML objects
An HTML report is a special object that is **not** run via `{r} pipeline_run()`. It is often indicated by `tag: html` and has a dedicated mapping key for the output location of the HTML file (this path does not need to vary per run, as the `RUN_TIMESTAMP` is used in the report output file name):

```{yaml example_dir_html, eval=FALSE}
html_report:
   type: object
  tag: html
  dir_html: output/report/
```
> Note: For different reports we advise using different folders. This is why the `dir_html` path is not defined in `file_paths.yaml`.

#### Mapping keys for `type: data`

Objects with `type: data` represent input coming from a datasource pipeline.
This type has one key that is specific to it:

- `dir_input`
  The file path to the data produced by the datasource pipeline.

```{yaml example_dir_input, eval = FALSE}
dat_osiris:
  type: data
  dir_input: /{TARGET_MOUNT}/EPI/DIS/DataPipelines/datasource-osiris/data/
```
> Note: We use the `{TARGET_MOUNT}` variable because the mount point differs between manual runs and cron jobs.


#### Mapping keys for `type: other`

Objects with type: other represent fixed, external files that do not change over time, in contrast to datasource files (`type: data`), where the most recent data (based on `RUN_TIMESTAMP`) is used automatically.

This type also uses a single key:
- `dir_input`
The file path to the static input file.

```{yaml example_type_other, eval = FALSE}
my_constant_xlsx:
  type: other
  dir_input: /rivm/EPI/DIS/FixedData/data_static/data.xlsx
```

> Note: If data is refreshed (e.g. daily, weekly, monthly), it should be defined as `type: data`.
Only truly constant inputs should use `type: other`.

## Config file `file_paths.yaml`
This config file stores the locations of pipeline scripts, outputs, and other shared resources.
It has two required keys (the index names on the left side of the `:`):

- `dir_scripts` – folder containing the scripts for the objects to run
- `dir_outputs` – folder where pipeline output is saved

Two other, optional keys commonly used in this file are:

- `dir_proj` – the project directory, often structured as `department/DataPipelines/repository_name`
- `dir_mnt` – the mount point where the project directory is available, often `/rivm/EPI`

Other useful shared file paths can also be stored in `file_paths.yaml`.
File paths that are specific to a single object should be stored in `object_definitions.yaml`.

Example `file_paths.yaml`:

```{yaml example_file_paths, eval = FALSE}
dir_outputs: output/data_objects/{RUN_TIMESTAMP}
dir_scripts: scripts/01_objects/{TAG}
dir_proj: 'DIS/DataPipelines/report-surveillance'
dir_mnt: "/rivm/EPI"
```

## Config file `object_relations.yaml`

This config file defines the input–output relations between all object types:

- `type: data` – used only as input (listed with a dash `-` in front)
- `type: other` – used only as input
- `type: object` – can appear both as input and as a key (the object being created)

Example:
```{yaml example_object_relations, eval = FALSE}
my_figure:
  - dat_osiris
  - my_constant_xlsx
```
This should be read as: `my_figure` **depends on** `dat_osiris` **and** `my_constant_xlsx`.

It is important to list **only** the inputs that are actually used in the script for that object, to keep the dependency graph accurate and minimal.

## Config file `object_schedules.yaml`

This optional config file can be used to define schedules for either tags or objects. A schedule defines
when an object should be run. If no schedules are defined, all objects are created. A schedule for an
object consists of three elements:
- `day`: The day that the object should be created (in English)
- `start`: The start time of the schedule window for the given day
- `end`: The end time of the schedule window for the given day

Objects and tags can have multiple schedules, either on the same day or different days. For example:

```{yaml example_object_schedules, eval = FALSE}
my_object:
  - day: tuesday
    start: "12:00"
    end: "12:30"
  - day: wednesday
    start: "12:00"
    end: "12:30"

my_tag:
  - day: monday
    start: "09:00"
    end: "10:30"
```

In this example, `my_object` will only run if the pipeline is run on tuesday or wednesday between
12:00 and 12:30. Objects with tag `my_tag` will only run if the pipeline is run on monday between
09:00 and 10:30.

> Note: objects with a schedule that are dependencies of other objects will still be created. E.g.,
if a pipeline is run outside the schedule of `my_object`, but `my_object` is needed to create a
different object, `my_object` will still be run, even though its schedule is not active.

# functions

## `pipeline_run()`

We already discussed `pipeline_init()`.
The next main function is `pipeline_run()`. Based on what is listed in `object_relations.yaml`, it:

1. **Loads input**
   All required inputs for each object are loaded, if they are not already in memory.

2. **Runs the scripts**
   The object script is executed. The script location is determined using:
   - `file_paths.yaml` (i.e. `dir_scripts`)
   - the object name (which should match the script filename).

3. **Caches and saves output**
   Caching and output saving are controlled by the settings in `object_definitions.yaml`
   (e.g. `skip_cache`, `cache_dir`, `output_formats`, `output_formats_custom`)
   and by the output locations (i.e. `dir_outputs`) defined in `file_paths.yaml`.

The order in which objects are run is determined from their dependencies in `object_relations.yaml`.

Two useful parameters of `pipeline_run()` are:

- `objects` – run a specific object or list of objects.
- `tags` – run all objects that share one or more tags.

Example:

```{r eval=FALSE}
# to only run objects + dependencies with these two tags
pipeline_run(tags = c("tag_A", "tag_B"))

# to only run one or multiple objects by name (+ dependencies)
pipeline_run(objects = "figure_A")
```

## `pipeline_vis()`

`pipeline_vis()` visualizes the Directed Acyclic Graph (DAG), i.e. a graph of
all dependencies between objects, based on `object_relations.yaml`.

This helps to:

- check whether all dependencies and tags are defined correctly
- see the order in which objects will be run
- identify shared inputs between different objects

```{r example_pipeline_vis, eval=TRUE, fig.width = 8, fig.height = 4.5}
pipeline_vis()
```

## `pipeline_import_for()`

HTML reports are not generated by `pipeline_run()`. Instead, they are typically created at the end of `report.R`.
However, the report still needs all data and objects that the HTML depends on.

For an HTML object with the name `html_report`, you can import all its dependencies with:

```{r eval = FALSE}
pipeline_import_for(data_assets = "html_report")
```

This loads all required inputs (as defined in `object_relations.yaml`) into memory, so they are available when you build the report.


## Getting started with your own pipeline project
To create your own report pipeline project using the `DARApipeline` template, you can start with the cookiecutter. Installation and usage instructions are available [here](
https://gitlab.rivm.nl/dara/wiki/-/wikis/Tutorial:-How-to-install-cookiecutter-and-start-a-new-EPI-pipeline)
